Clean data

### 2023-05-16 analysis/02_clean-data.Rmd
Cleaned and restructured 02_clean-data

## Design matrix

Select a dummy or index-coded version of the design.

```{r transform-design}
dummy_coded <- 0 # Dummy-coded design and associated interactions.
index_coded <- 1 # Index-coded design and associated interactions.

if (dummy_coded == 1) {
  # Load the dummy design.
  dummy_design <- read_csv(here::here("data", "derived_data", "dummy_design.csv")) %>%
    select(-`...1`) %>%
    rename(
      version = question_set,
      task = question,
      alt = choice
    )

@@

  # Reintroduce org1 and add the outside option coded as all zeros.
  for (n in 1:dim(dummy_X)[1]) {
    # Filter for respondent n.
    dummy_X_n <- dummy_design %>% filter(version == data$version[n])
    interactions_X_n <- dummy_interactions_design %>% filter(version == data$version[n])
    for (s in 1:dim(dummy_X)[2]) {
      # Filter for task s and include placeholder for org1.
      dummy_X_s <- dummy_X_n %>%
        filter(task == s) %>%
        mutate(org1 = 0) %>%
        select(org1, org2:gov_rel3)
      interactions_X_s <- interactions_X_n %>%
        filter(task == s) %>%
        mutate(org1 = 0) %>%
        select(org1, org2:transp_smalldonors_refugees_crackdown) # NOTE DIMENSIONS OF INTERACTIONS HARD-CODED.
      for (p in 1:(dim(dummy_X)[3]-1)) {
        # Filter for task s and alt p orgs.
        dummy_X_p <- dummy_X_n %>%
          filter(task == s, alt == p) %>%
          select(org2:org4) %>%
          as_vector()
        interactions_X_p <- interactions_X_n %>%
          filter(task == s, alt == p) %>%
          select(org2:org4) %>%
          as_vector()
        # Fill in org1 to complete org intercept.
        dummy_X_s[p, "org1"] <- ifelse(sum(dummy_X_p) == 1, 0, 1)
        interactions_X_s[p, "org1"] <- ifelse(sum(interactions_X_p) == 1, 0, 1)
      }
      # Save modified design, including outside option.
      dummy_X[n, s,,] <- rbind(dummy_X_s, rep(0, ncol(dummy_X_s))) %>% as.matrix()
      dummy_interactions_X[n, s,,] <- rbind(interactions_X_s, rep(0, ncol(interactions_X_s))) %>% as.matrix()
    }
  }
}
if (index_coded == 1) {
  # Load the index-coded design.
  index_design <- read_csv(here::here("data", "derived_data", "index_design.csv")) %>%
    select(-`...1`) %>%
    rename(
      version = question_set,
      task = question,
      alt = choice
    )

@@

  # Add the outside option coded as all zeros (org1 and former reference levels already included).
  for (n in 1:dim(index_X)[1]) {
    # Filter for respondent n.
    index_X_n <- index_design %>% filter(version == data$version[n])
    interactions_X_n <- index_interactions_design %>% filter(version == data$version[n])
    for (s in 1:dim(index_X)[2]) {
      # Filter for task s.
      index_X_s <- index_X_n %>%
        filter(task == s) %>%
        select(org1:gov_rel3)
      interactions_X_s <- interactions_X_n %>%
        filter(task == s) %>%
        select(org1:transp_smalldonors_refugees_crackdown) # NOTE DIMENSIONS OF INTERACTIONS HARD-CODED.
      # Save modified design, including outside option.
      index_X[n, s,,] <- rbind(index_X_s, rep(0, ncol(index_X_s))) %>% as.matrix()
      index_interactions_X[n, s,,] <- rbind(interactions_X_s, rep(0, ncol(interactions_X_s))) %>% as.matrix()
    }
  }
}
```

@@

```{r save-clean-data}
write_rds(data, here("data", "derived_data", "final_data.rds"))
if (dummy_coded == 1) {
  write_rds(dummy_X, here("data", "derived_data", "final_design.rds"))
  write_rds(dummy_interactions_X, here("data", "derived_data", "final_interactions.rds"))
}
if (index_coded == 1) {
  write_rds(index_X, here("data", "derived_data", "final_design.rds"))
  write_rds(index_interactions_X, here("data", "derived_data", "final_interactions.rds"))
}
```

>>>>

## Dummy-coded design matrix

Transform the dummy-coded design matrix conditioned on what version each respondent was assigned to, including the brand intercept and the outside option.

```{r transform-dummy-design}
# Load the dummy design.
dummy_design <- read_csv(here::here("data", "derived_data", "dummy_design.csv")) %>%
  select(-`...1`) %>%
  rename(
    version = question_set,
    task = question,
    alt = choice
  )

@@

# Reintroduce org1 and add the outside option coded as all zeros.
for (n in 1:dim(dummy_X)[1]) {
  # Filter for respondent n.
  dummy_X_n <- dummy_design %>% filter(version == data$version[n])
  interactions_X_n <- dummy_interactions_design %>% filter(version == data$version[n])
  for (s in 1:dim(dummy_X)[2]) {
    # Filter for task s and include placeholder for org1.
    dummy_X_s <- dummy_X_n %>%
      filter(task == s) %>%
      mutate(org1 = 0) %>%
      select(org1, org2:gov_rel3)
    interactions_X_s <- interactions_X_n %>%
      filter(task == s) %>%
      mutate(org1 = 0) %>%
      select(org1, which(names(interactions_X_n) == "org2"):ncol(interactions_X_n))
    for (p in 1:(dim(dummy_X)[3]-1)) {
      # Filter for task s and alt p orgs.
      dummy_X_p <- dummy_X_n %>%
        filter(task == s, alt == p) %>%
        select(org2:org4) %>%
        as_vector()
      interactions_X_p <- interactions_X_n %>%
        filter(task == s, alt == p) %>%
        select(org2:org4) %>%
        as_vector()
      # Fill in org1 to complete org intercept.
      dummy_X_s[p, "org1"] <- ifelse(sum(dummy_X_p) == 1, 0, 1)
      interactions_X_s[p, "org1"] <- ifelse(sum(interactions_X_p) == 1, 0, 1)
    }
    # Save modified design, including outside option.
    dummy_X[n, s,,] <- rbind(dummy_X_s, rep(0, ncol(dummy_X_s))) %>% as.matrix()
    dummy_interactions_X[n, s,,] <- rbind(interactions_X_s, rep(0, ncol(interactions_X_s))) %>% as.matrix()
  }
}
```

## Index-coded experimental alternative

```{r transform-index-design}
################################################################################
# This is an experimental alternative to the dummy-coded design.
################################################################################

# Load the index-coded design.
index_design <- read_csv(here::here("data", "derived_data", "index_design.csv")) %>%
  select(-`...1`) %>%
  rename(
    version = question_set,
    task = question,
    alt = choice
  )

@@

# Add the outside option coded as all zeros (org1 and former reference levels already included).
for (n in 1:dim(index_X)[1]) {
  # Filter for respondent n.
  index_X_n <- index_design %>% filter(version == data$version[n])
  interactions_X_n <- index_interactions_design %>% filter(version == data$version[n])
  for (s in 1:dim(index_X)[2]) {
    # Filter for task s.
    index_X_s <- index_X_n %>%
      filter(task == s) %>%
      select(org1:gov_rel3)
    interactions_X_s <- interactions_X_n %>%
      filter(task == s) %>%
      select(which(names(interactions_X_n) == "org1"):ncol(interactions_X_n))
    # Save modified design, including outside option.
    index_X[n, s,,] <- rbind(index_X_s, rep(0, ncol(index_X_s))) %>% as.matrix()
    index_interactions_X[n, s,,] <- rbind(interactions_X_s, rep(0, ncol(interactions_X_s))) %>% as.matrix()
  }
}
```

## Flat MNL designs experimental alternative

```{r transform-flat-designs}
################################################################################
# This is an experimental designs for a flat MNL.
# We already have the data structured. We just need to collapse respondent
# repeat observations into an observation dimension.
################################################################################

# Create the arrays for a design with a brand intercept.
flat_dummy_X <- array(
  data = NA,
  dim = c(
    nrow(data) * max(dummy_design$task), # Number of observations.
    (max(dummy_design$alt) + 1),         # Number of product alternatives per observation.
    (ncol(dummy_design) - 3 + 1)         # Number of (estimable) attribute levels.
  )
)

# Create the arrays for a design with a brand intercept and interactions.
flat_dummy_interactions_X <- array(
  data = NA,
  dim = c(
    nrow(data) * max(dummy_interactions_design$task), # Number of respondents.
    (max(dummy_interactions_design$alt) + 1),         # Number of product alternatives per observation.
    (ncol(dummy_interactions_design) - 3 + 1)         # Number of (estimable) attribute levels.
  )
)

# Create the arrays for the design, including a brand intercept.
flat_index_X <- array(
  data = NA,
  dim = c(
    nrow(data) * max(index_design$task), # Number of observations.
    (max(index_design$alt) + 1),         # Number of product alternatives per observation.
    (ncol(index_design) - 3)             # Number of (estimable) attribute levels.
  )
)

# Create the arrays for a design, including a brand intercept and interactions.
flat_index_interactions_X <- array(
  data = NA,
  dim = c(
    nrow(data) * max(index_interactions_design$task), # Number of respondents.
    (max(index_interactions_design$alt) + 1),         # Number of product alternatives per observation.
    (ncol(index_interactions_design) - 3)             # Number of (estimable) attribute levels.
  )
)

# Restructure design matrices.
n <- 0
for (r in 1:dim(dummy_X)[1]) {
  for (s in 1:dim(dummy_X)[2]) {
    n <- n + 1
    flat_dummy_X[n,,] <- dummy_X[r, s,,]
    flat_dummy_interactions_X[n,,] <- dummy_interactions_X[r, s,,]
    flat_index_X[n,,] <- index_X[r, s,,]
    flat_index_interactions_X[n,,] <- index_interactions_X[r, s,,]
  }
}
```

@@

```{r save-clean-data}
write_rds(data, here("data", "derived_data", "data.rds"))
write_rds(dummy_X, here("data", "derived_data", "hmnl_dummy_design.rds"))
write_rds(dummy_interactions_X, here("data", "derived_data", "hmnl_dummy_interactions.rds"))

################################################################################
# Save the experimental alternative to the dummy-coded design.
################################################################################

write_rds(index_X, here("data", "derived_data", "hmnl_index_design.rds"))
write_rds(index_interactions_X, here("data", "derived_data", "hmnl_index_interactions.rds"))

################################################################################
# Save the experimental designs for a flat MNL.
################################################################################

write_rds(flat_dummy_X, here("data", "derived_data", "mnl_dummy_design.rds"))
write_rds(flat_dummy_interactions_X, here("data", "derived_data", "mnl_dummy_interactions.rds"))
write_rds(flat_index_X, here("data", "derived_data", "mnl_index_design.rds"))
write_rds(flat_index_interactions_X, here("data", "derived_data", "mnl_index_interactions.rds"))
```
